

```{r}
library(lme4)
library(tidyverse)
library(car)
library(ggplot2)
library(broom)
library(emmeans)
library(ggpubr)
library(caret)
library(knitr)
library(glue)
library(ggplot2)
library(ggpmisc)
library(ggrepel)

```


```{r}
# load in data

obs_df <- read.csv("Folder3-Gene_Shuffle_Predictions/prot_expression_ratio_data.csv")

pred_df <- read.csv("Folder3-Gene_Shuffle_Predictions/RAM_mean_predictions.csv")
pred_std <- read.csv("Folder3-Gene_Shuffle_Predictions/RAM_std_predictions.csv")

```



```{r}
# pivot data table to better format
long_pred <- pred_df %>%
  pivot_longer(cols = N:L, names_to = "Variable", values_to = "Predicted_expression")

```

```{r}
# pivot data table to better format
long_obs <- obs_df %>%
  pivot_longer(cols = N:L, names_to = "Variable", values_to = "Observed_expression")
long_obs$Predicted_expression <- long_pred$Predicted_expression
long_obs$Gene <- factor(long_obs$Variable, levels = c("N", "P", "M", "G", "L")) 

```



```{r}
# check model assumptions
mod_gn <- lm(Predicted_expression~Observed_expression, data=long_obs)
performance::check_model(mod_gn, check="linearity")
performance::check_model(mod_gn, check="homogeneity")
performance::check_model(mod_gn, check="qq")
tidy(mod_gn, conf.int=TRUE, conf.level=0.95)
# there are some potential minor violations. mainly appear to be due to the one outlier
```

```{r}
0.8478145 - 0.696286017
```

```{r}
summary(mod_gn)
```



```{r}
long_std <- pred_std %>%
  pivot_longer(cols = N:L, names_to = "Variable", values_to = "Predicted_sd")

long_obs$Predicted_sd <- long_std$Predicted_sd

g_df <- ggplot(long_obs, aes(x = Observed_expression, y = Predicted_expression, color = Gene, shape = Gene)) +
  theme_bw() +
  geom_point(size = 3) +
geom_errorbar(aes(ymin = Predicted_expression - 2 * Predicted_sd,
                  ymax = Predicted_expression + 2 * Predicted_sd),
              width = 0.02,
              size = 0.8,
              alpha = 0.8) +
  xlim(0, 0.6) +
  ylim(0, 0.45) +
  xlab("Observed Relative Expression (Protein Level)") +
  ylab("Predicted Relative\nExpression (Transcript Level)") +
  geom_smooth(aes(group = 1), 
              method = "lm", se = FALSE, color = "black", size = 1) + 
  theme(
    axis.text = element_text(size = 16),           
    axis.title = element_text(size = 20, face = "bold"),
    legend.title = element_text(size = 14, face = "bold"), 
    legend.text = element_text(size = 12),
    legend.position = "bottom"
  ) + 
  stat_poly_eq(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~"), group = 1),
               label.x.npc = "right", label.y.npc = 0.9, 
               formula = y ~ x, parse = TRUE, size = 6) +
  guides(color = guide_legend(override.aes = list(size = 4)))

ggsave("Figures/Fig6/all_RAM_plot.svg", plot = g_df, device = "svg", width = 10, height = 6, dpi = 300)

g_df


```
gene_cols <- ggplot_build(g_df)$plot$scales$get_scales("colour")$palette(5)
names(gene_cols) <- levels(long_obs$Gene)
gene_cols



```{r}

# load in data

obs_df <- read.csv("Folder3-Gene_Shuffle_Predictions/prot_expression_ratio_data.csv")

pred_df <- read.csv("Folder3-Gene_Shuffle_Predictions/ss_mean_predictions.csv")
pred_std <- read.csv("Folder3-Gene_Shuffle_Predictions/ss_std_predictions.csv")
```



```{r}
# pivot data table to better format
long_pred <- pred_df %>%
  pivot_longer(cols = N:L, names_to = "Variable", values_to = "Predicted_expression")

```

```{r}
# pivot data table to better format
long_obs <- obs_df %>%
  pivot_longer(cols = N:L, names_to = "Variable", values_to = "Observed_expression")
long_obs$Predicted_expression <- long_pred$Predicted_expression
long_obs$Gene <- factor(long_obs$Variable, levels = c("N", "P", "M", "G", "L")) 

```




```{r}
# check model assumptions
mod_gn <- lm(Predicted_expression~Observed_expression, data=long_obs)
performance::check_model(mod_gn, check="linearity")
performance::check_model(mod_gn, check="homogeneity")
performance::check_model(mod_gn, check="qq")
tidy(mod_gn, conf.int=TRUE, conf.level=0.95)
# there are some potential minor violations. mainly appear to be due to the one outlier
```

```{r}
summary(mod_gn)
```


```{r}
0.87153194 - 0.71055594
```



```{r}
long_std <- pred_std %>%
  pivot_longer(cols = N:L, names_to = "Variable", values_to = "Predicted_sd")

long_obs$Predicted_sd <- long_std$Predicted_sd

g_df <- ggplot(long_obs, aes(x = Observed_expression, y = Predicted_expression, color = Gene, shape = Gene)) +
  theme_bw() +
  geom_point(size = 3) +
geom_errorbar(aes(ymin = Predicted_expression - 2 * Predicted_sd,
                  ymax = Predicted_expression + 2 * Predicted_sd),
              width = 0.02,
              size = 0.8,
              alpha = 0.8) +
  xlim(0, 0.6) +
  ylim(0, 0.45) +
  xlab("Observed Relative Expression (Protein Level)") +
  ylab("Predicted Relative\nExpression (Transcript Level)") +
  geom_smooth(aes(group = 1), 
              method = "lm", se = FALSE, color = "black", size = 1) + 
  theme(
    axis.text = element_text(size = 16),           
    axis.title = element_text(size = 20, face = "bold"),
    legend.title = element_text(size = 14, face = "bold"), 
    legend.text = element_text(size = 12),
    legend.position = "bottom"
  ) + 
  stat_poly_eq(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~"), group = 1),
               label.x.npc = "right", label.y.npc = 0.9, 
               formula = y ~ x, parse = TRUE, size = 6) +
  guides(color = guide_legend(override.aes = list(size = 4)))

# Save the plot in high resolution
ggsave("Figures/Fig6/all_ss_plot.svg", plot = g_df, device = "svg", width = 10, height = 6, dpi = 300)

# Display the plot
g_df


```

```{r}
library(ggplot2)
library(dplyr)
library(scales)

df <- read.csv("Folder3-Gene_Shuffle_Predictions/model_absolute_errors.csv")

genes <- c("N","P","M","G","L")
df$Model <- factor(df$Model, levels = c("RAM", "Stop-Start"))
df$Gene  <- factor(df$Gene,  levels = genes)

gene_cols <- setNames(hue_pal()(length(genes)), genes)

gene_shapes <- c(
  N = 16,  
  P = 17,  
  M = 15,
  G = 3, 
  L = 7  
)

summary_df <- df %>%
  group_by(Model) %>%
  summarise(
    mean = mean(Absolute_Error),
    se   = sd(Absolute_Error)/sqrt(n()),
    ci   = 1.96 * se,
    .groups = "drop"
  )

mae_ram <- mean(df$Absolute_Error[df$Model == "RAM"])
mae_ss  <- mean(df$Absolute_Error[df$Model == "Stop-Start"])
rel_improve <- abs(mae_ram - mae_ss) / mae_ram * 100

cat(sprintf("Overall MAE RAM: %g\n", mae_ram))
cat(sprintf("Overall MAE Stop-Start: %g\n", mae_ss))
cat(sprintf("Relative improvement (RAM->SS): %.2f%%\n\n", rel_improve))

per_gene <- df %>%
  group_by(Gene, Model) %>%
  summarise(
    mean = mean(Absolute_Error),
    sd   = sd(Absolute_Error),
    .groups = "drop"
  ) %>%
  mutate(`mean_pm_2sd` = sprintf("%0.3g ± %0.3g", mean, 2*sd))

print(per_gene)

n_each <- sum(df$Model == "RAM")
stopifnot(sum(df$Model == "Stop-Start") == n_each)

err_ram <- df$Absolute_Error[df$Model == "RAM"]
err_ss  <- df$Absolute_Error[df$Model == "Stop-Start"]

tt <- t.test(err_ram, err_ss, paired = TRUE)
print(tt)

metrics_text <- sprintf(
  "MAE RAM: %.3g\nMAE SS : %.3g\nPaired t-test p: %.2g",
  mae_ram, mae_ss, tt$p.value
)

mae_ram <- mean(df$Absolute_Error[df$Model == "RAM"])
mae_ss  <- mean(df$Absolute_Error[df$Model == "Stop-Start"])

pct_diff <- (mae_ss - mae_ram) / mae_ram * 100

metrics_text <- sprintf(
  "MAE RAM: %.3g\nMAE SS : %.3g",
  mae_ram, mae_ss, pct_diff
)

p <- ggplot(df, aes(x = Model, y = Absolute_Error)) +
  geom_violin(
    aes(group = Model),
    trim = FALSE,
    fill = "grey90",
    color = "grey35",
    linewidth = 0.8
  ) + ylim(0, 0.25) +
  geom_jitter(
    aes(color = Gene, shape = Gene),
    width = 0.30,
    size = 3.0,    
    alpha = 0.75,
    stroke = 0.8  
  ) +
  geom_errorbar(
    data = summary_df,
    aes(x = Model, y = mean, ymin = mean - ci, ymax = mean + ci),
    width = 0.12,
    linewidth = 1.0,
    inherit.aes = FALSE,
    color = "black"
  ) +
  geom_point(
    data = summary_df,
    aes(x = Model, y = mean),
    size = 3.4,
    inherit.aes = FALSE,
    color = "black"
  ) +
  scale_color_manual(values = gene_cols) +
  scale_shape_manual(values = gene_shapes) +
  labs(y = "Absolute Error", x = NULL, color = "Gene", shape = "Gene") +
  annotate(
    "label",
    x = 0.1, y = 0.245,        
    label = metrics_text,
    hjust = 0, vjust = 1,
    size = 5,
    label.size = 0.25
  ) +
  coord_cartesian(ylim = c(0, 0.25)) +
  theme_bw() +
  theme(
    axis.text = element_text(size = 16),
    axis.title = element_text(size = 20, face = "bold"),
    legend.title = element_text(size = 14, face = "bold"),
    legend.text = element_text(size = 12),
    legend.position = "bottom",
    plot.margin = margin(10, 10, 10, 10)
  ) +
  guides(
    color = guide_legend(override.aes = list(size = 4)),
    shape = guide_legend(override.aes = list(size = 4))
  )
ggsave("Figures/Fig6/all_barplot.svg", plot = p, device = "svg", width = 10, height = 6, dpi = 300)

```





